---
title: "DAGPD: Niche-Dependent Functional Genes in Diverse Domestic Animals"
author: "Yuxuan Huang"
format: 
  html:
    embed-resources: true
---

# Niche-Dependent Functional genes in Diverse Domestic Animals

## volcano plot of AMGs relate to Animal hosts

```{r,warning=FALSE,message=FALSE}
setwd("~/DAGPD/Scripts/")
.libPaths("/home/lsx/R/x86_64-pc-linux-gnu-library/4.4/")
source("~/DAGPD/Scripts/theme_llm.R")

require(ggplot2)

library(ggrepel)
library(tidyverse)

library(viridis)  # color gradation
gene_hits = read.csv("../Pre-processed_Files/Animal_relate_AMG2.csv", stringsAsFactors=FALSE, header=TRUE)
gene_hits$logor <- as.numeric(gene_hits$logor)

gene_hits <- filter(gene_hits,gene_hits$adj.p != 0)

theme_set(theme_llm(base_size = 12,legend = "right"))
ggplot(gene_hits, aes(x=logor, y=logpadj, colour=logpadj, size=abs(logor), label=label)) +
  geom_point(alpha=0.5) +
  geom_text_repel(
    data = subset(gene_hits, label != ""),  # sel points
    aes(label = label),
    box.padding = 1,          # padding between point and label
    segment.alpha = 0.5,
    min.segment.length = 0.2,
    max.overlaps = 52,         # overlap of labels
    size = 3,
    segment.color = "grey50",    # line color
    color='black')+
  geom_hline(yintercept = 1.3,           # line position
             linetype = "dashed",      # linetype
             color = "grey50",            # line color
             linewidth = 1)+
  xlim(c(-50,200))+
  ylim(c(-10,160))+
  xlab(expression(Log[10]~"Odds Ratio"))+
  ylab(expression(-Log[10] ~ "FDR") )+
  scale_color_gradient(limits = c(0, 160))+
  ggtitle("Animal Hosts")


```

## volcano plot of AMGs relate to Bacterial hosts

```{r,warning=FALSE,message=FALSE}
setwd("~/DAGPD/Scripts/")
gene_hits_bac = read.csv("../Pre-processed_Files/Bacterial_relate_AMG2.csv", stringsAsFactors=FALSE, header=TRUE)
gene_hits_bac$logor <- as.numeric(gene_hits_bac$logor)
gene_hits_bac$logpadj <- as.numeric(gene_hits_bac$logpadj)
gene_hits_bac <- filter(gene_hits_bac,gene_hits_bac$adj.p != 0)

theme_set(theme_llm(base_size = 12,legend = "right"))
ggplot(gene_hits_bac, aes(x=logor, y=logpadj, colour=logpadj, size=abs(logor), label=label)) +
  geom_point(alpha=0.5) +
  geom_text_repel(
    data = subset(gene_hits_bac, label != ""),  # 
    aes(label = label),
    box.padding = 0.1,          # 
    segment.alpha = 0.5,
    min.segment.length = 0.2,
    max.overlaps = 50,         # 
    size = 3,
    segment.color = "grey50",    # 
    color='black')+
  geom_hline(yintercept = 1.3,           # 
             linetype = "dashed",      # 
             color = "grey50",            # 
             linewidth = 1)+
  xlim(c(-10,40))+
  ylim(c(0,8))+
  xlab(expression(Log[10]~"Odds Ratio"))+
  ylab(expression(-Log[10] ~ "FDR") )+
  scale_color_gradient(limits = c(0, 10))+
   ggtitle("Bacterial Hosts")
```
## Diversity of AMGs from different Animals and Locations

### PcoA analysis of AMGs
```{r PcoA,warning=FALSE,message=FALSE}

#points = as.data.frame(pcoa$points) 
points <- read.csv("../Pre-processed_Files/AMG_PcoA_bray_point.csv",header = T,row.names = 1)
Group1 <- read.table("../Pre-processed_Files/AMG_Group1.txt",header = T,sep = "\t")
Group1 <- column_to_rownames(Group1,"sample_id")
points <- merge(points,Group1, by = "row.names")
colnames(points) <- c("sample_id","V1","V2","V3","location","Species","Source")
#points$Source=Group1$V2
#points$location=Group1$V3
#points$Class=Group1$V4
# wascores()
#spe.wa <- wascores(pcoa$points, spe)
load("../Pre-processed_Files/AMG_pcoa.RData")
eig = pcoa$eig

theme_set(theme_llm(base_size = 12,margin=FALSE))
ggplot(points, aes(x=V1, y=V2, colour=Source))+geom_point(alpha=0.7, size=1)+
  scale_color_manual(
    values = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD"))+
  #scale_shape_manual(
  #  values = c("Asia"= 16,"Europe"=17,"North America"=18,"Oceania"=19)
#  )+
  stat_ellipse()+
  labs(x = paste("PCoA 1 (", format(100*pcoa$eig[1] / sum(pcoa$eig), digits = 4), "%)",sep = ""), 
       y = paste("PCoA 2 (", format(100*pcoa$eig[2] / sum(pcoa$eig), digits = 4), "%)",sep = ""))

ggplot(points, aes(x=V1, y=V2, colour=location))+geom_point(alpha=0.7, size=1)+
  scale_color_manual(
    values = c("Asia"="#FB8072","Europe"="#BEBADA","North America"="#FDB462","Oceania"="#80B1D3"))+
  stat_ellipse()+
  labs(x = paste("PCoA 1 (", format(100*pcoa$eig[1] / sum(pcoa$eig), digits = 4), "%)",sep = ""), 
       y = paste("PCoA 2 (", format(100*pcoa$eig[2] / sum(pcoa$eig), digits = 4), "%)",sep = ""))

```

```{r NMDC,message=FALSE,warning=FALSE}


#NMDSpoints1=as.data.frame(NMDS1$points)
NMDSpoints1 <- read.csv("../Pre-processed_Files/AMG_NMDSpoints1.csv",header = T,row.names = 1)
head(NMDSpoints1)


#Species1=NMDS1$species
#head(Species1)



library(tidyverse)
Group1 <- read.table("../Pre-processed_Files/AMG_Group1.txt",header = T,sep = "\t")
Group1 <- column_to_rownames(Group1,"sample_id")
NMDSpoints1 <- merge(NMDSpoints1,Group1, by = "row.names")
colnames(NMDSpoints1) <- c("sample_id","NDS1","NDS2","location","Species","Source")
theme_set(theme_llm(base_size = 12,margin=FALSE))
#NMDS AMG
ggplot(NMDSpoints1, aes(x=NDS1, y=NDS2, colour=Source))+geom_point(alpha=0.7, size=1)+
  scale_color_manual(values = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD"))+
  stat_ellipse()+
  xlab("NMDS1")+
  ylab("NMDS2")

ggplot(NMDSpoints1, aes(x=NDS1, y=NDS2, colour=location))+geom_point(alpha=0.7, size=1)+
  scale_color_manual(values = c("Asia"="#FB8072","Europe"="#BEBADA","North America"="#FDB462","Oceania"="#80B1D3"))+
  stat_ellipse()+
  xlab("NMDS1")+
  ylab("NMDS2")

```


## Bubble plot of KEGG enrichment analysis

```{r KEGG,warning=FALSE,message=FALSE}
setwd("~/DAGPD/Scripts/")
data1 <- read.csv("../Pre-processed_Files/AMG_KEGG_ClassI.csv",header = T)
# 绘制气泡图
theme_set(theme_llm(base_size = 12,legend = "right"))
data1$Category <- factor(data1$Category,levels = rev(c("Carbohydrate metabolism",
                                                   "Glycan biosynthesis and metabolism",
                                                   "Metabolism of terpenoids and polyketides",
                                                   "Metabolism of cofactors and vitamins",
                                                   "Energy metabolism",
                                                   "Lipid metabolism",
                                                   "Amino acid metabolism",
                                                   "Metabolism of other amino acids",
                                                   "Protein families: metabolism",
                                                   "Cellular community",
                                                   "Signal transduction",
                                                   "Translation"
                                                   )))
category_colors <- data.frame(
  Category = rev(c("Carbohydrate metabolism",
                   "Glycan biosynthesis and metabolism",
                   "Metabolism of terpenoids and polyketides",
                   "Metabolism of cofactors and vitamins",
                   "Energy metabolism",
                   "Lipid metabolism",
                   "Amino acid metabolism",
                   "Metabolism of other amino acids",
                   "Protein families: metabolism",
                   "Cellular community",
                   "Signal transduction",
                   "Translation"
  )),
  Color = c("#A5A5A5","#A5A5A5","#A5A5A5","#A5A5A5","#E4B8ED","#E4B8ED","#F8B535","#4E76BB","#9767C7","#C8EDF7","#AEE57C","#8DD3C7")
)

ggplot() +
  geom_tile(data = category_colors, aes(x = 0, y = Category, fill = Color), width = 0.2, height = 1)+
  scale_fill_identity() +
  geom_point(data=data1, aes(x = Group, y = Category, size = Enrichfactor, color = logp))+ #geom_point(data = data, aes(x = Group, y = Category, size = Size, color = Color), alpha = 0.7) +
  geom_point(alpha = 0.8) +  # transparent
  scale_x_discrete(expand = expansion(mult = c(0, 0.5))) +
  scale_size_continuous(name = "Enrichment Factor", 
                        range = c(1, 6),  # bubble size
                        breaks = seq(1, 5, by = 1)) +  # 
  scale_color_viridis(name = "-log10(p-value)", 
                      option = "viridis",  # color
                      direction = -1) +  # color direction
  labs(x = "Group", y = "Category", 
       title = "Bubble Chart by Animals")+theme(aspect.ratio=2)+
  theme(
    axis.text.y = element_text(margin = margin(r = 0))  # right margin
  )


```

```{r,warning=FALSE,message=FALSE}
# class2
data2 <- read.csv("../Pre-processed_Files/AMG_KEGG_Class2.csv",header = T)

data2$Category <- factor(data2$Category,levels = rev(c("Glycolysis / Gluconeogenesis",
                                                   "Pentose phosphate pathway",
                                                   "Pyruvate metabolism",
                                                   "Starch and sucrose metabolism",
                                                   "Benzoate degradation",
                                                   "Fructose and mannose metabolism",
                                                   "Peptidoglycan biosynthesis and degradation proteins",
                                                   "Polyketide sugar unit biosynthesis",
                                                   "One carbon pool by folate",
                                                   "Vitamin B6 metabolism",
                                                   "Glycerophospholipid metabolism",
                                                   "Phenylalanine, tyrosine and tryptophan biosynthesis"
                                                   )))
category_colors <- data.frame(
  Category = c("Glycolysis / Gluconeogenesis",
               "Pentose phosphate pathway",
               "Pyruvate metabolism",
               "Starch and sucrose metabolism",
               "Benzoate degradation",
               "Fructose and mannose metabolism",
               "Peptidoglycan biosynthesis and degradation proteins",
               "Polyketide sugar unit biosynthesis",
               "One carbon pool by folate",
               "Vitamin B6 metabolism",
               "Glycerophospholipid metabolism",
               "Phenylalanine, tyrosine and tryptophan biosynthesis"
  ),
  Color = c("#8DD3C7","#8DD3C7","#8DD3C7","#8DD3C7","#8DD3C7","#8DD3C7","#AEE57C","#C8EDF7","#9767C7","#9767C7","#F8B535","#E4B8ED"))


theme_set(theme_llm(base_size = 12,legend = "left"))
ggplot() +
  geom_tile(data = category_colors, aes(x = 4, y = Category, fill = Color), width = 0.2, height = 1)+
  scale_fill_identity() +
  geom_point(data=data2, aes(x = Group, y = Category, size = Enrichfactor, color = logp))+ #geom_point(data = data, aes(x = Group, y = Category, size = Size, color = Color), alpha = 0.7) +
  geom_point(alpha = 0.8) +  # 
  scale_x_discrete(expand = expansion(mult = c(0.2, 0))) +
  scale_y_discrete(position = "right") +
  scale_size_continuous(name = "Enrichment Factor", 
                        range = c(1, 4.5),  # 
                        breaks = seq(1, 5, by = 1))+  # 
  scale_color_viridis(name = "-log10(p-value)", 
                      option = "viridis",  # 
                      limits = c(0, 16),
                      direction = -1) +  # 
  labs(x = "Group", y = "Category", 
       title = "Bubble Chart by Animals subclass")+theme(aspect.ratio=2)+
  theme(
    axis.text.y = element_text(margin = margin(r = 0))  # 
  )



```



## A bar plot showing alpha diversity of CAZy in phages among domestic animals

```{r,warning=FALSE,message=FALSE}

library(tidyverse)
source("./theme_llm.R")

cazy_alpha <- read.table("../Pre-processed_Files/AMG_cazy.txt",header = T)
cazy_alpha$Category <- factor(cazy_alpha$Category,
                              levels = rev(unique(cazy_alpha$Category)))
theme_set(theme_llm(base_size = 16,legend = "right"))
ggplot()+geom_col(data=cazy_alpha,mapping = aes(x=Sources,y=Counts,fill=Category),
                  width=0.5)+scale_fill_brewer(palette = "Set3")+
  ylab("Cazymes Diversity")+
  xlab("Animal Hosts")+
  ylim(0,810)
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r,warning=FALSE,message=FALSE}
# cazy <- read.table("../Pre-processed_Files/AMG_cazy_relative.txt",header = T)
# cazy$Category <- factor(cazy$Category,levels = rev(unique(cazy$Category)))
# theme_set(theme_llm(base_size = 16,legend = "right"))
# ggplot()+geom_col(data=cazy,mapping = aes(x=Sources,y=Counts/Ratio,fill=Category),
#                   width=0.5)+scale_fill_brewer(palette = "Set3")+
#   ylab("Cazymes per 1000 genes")+
#   xlab("Animal Hosts")+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## A bar plot showing COG annotation in different domestic animals
```{r,message=FALSE,echo=FALSE}
COG <- read.table("../Pre-processed_Files/COG_res.txt",header = T)
ggplot(COG, aes(x=COG, y=log(number,2), fill=Species)) + 
  geom_bar(stat = 'identity',position="dodge",alpha = 0.7)+
  labs(y = expression(paste("Number of Proteins (",x10^3, ")")),x=NULL)+
   scale_color_manual(values  = c("Chicken"="#FFBC00",
                                 "Pig"="#E588A6",
                                 "Ruminant"="#0AC5AD"),
                     aesthetics = c("fill"))
```
