---
title: "DAGPD: The coevolution and coadaptation of crAss-like phages with bacteria and animal hosts"
author: "Yuxuan Huang"
format: 
  html:
    embed-resources: true
---

# Coevolution and coadaptation with Prevotella and animals

## GC content of crAss-like phages alpha group

```{r,warning=FALSE,message=FALSE}

library(tidyverse)
source("./theme_llm.R")
theme_set(theme_llm(base_size = 16,legend = "none"))
# alpha group

GC <- read.table("../Pre-processed_Files/Alpha_GC.txt",header = T)
GC$Sources <- factor(GC$Sources,levels = c("Chicken", "Pig","Ruminant", "Human"))
library(ggsignif)
comparisons <- list(c("Pig", "Ruminant"),c("Ruminant", "Human"),c("Pig","Human"),
                    c("Chicken", "Pig"), c("Chicken", "Ruminant"),c("Chicken", "Human"))
ggplot(GC, aes(x=Sources, y=GC, color=Sources)) +
  geom_boxplot(size=0.5) +
   scale_color_manual(values  = c("Chicken"="#FFBC00","Pig"="#E588A6",
                                  "Ruminant"="#0AC5AD","Human"="#3FAAE2"))+
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_jitter(aes(color=Sources), size=2, alpha=0.7, width = 0.2)+
   geom_signif(
    comparisons = comparisons, test = "wilcox.test",
    map_signif_level = F, textsize = 4,
    step_increase = 0.1,color="black"
  ) +
   ylab("crAssphage GC (%)")+
   xlab("")+
   theme(legend.title=element_blank())+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## GC content of crAss-like phages beta group
```{r,warning=FALSE,message=FALSE}
# beta group
theme_set(theme_llm(base_size = 16,legend = "none"))
GC <- read.csv("../Pre-processed_Files/crass_beta_gc.csv",header = T)
GC$Sources <- factor(GC$Sources,levels = c("Chicken", "Pig","Ruminant", "Human"))
library(ggsignif)
comparisons <- list(c("Pig", "Ruminant"),c("Ruminant", "Human"),c("Pig","Human"),
                    c("Chicken", "Pig"), c("Chicken", "Ruminant"),c("Chicken", "Human"))
ggplot(GC, aes(x=Sources, y=GC, color=Sources)) +
  geom_boxplot(size=0.5) +
   scale_color_manual(values  = c("Chicken"="#FFBC00","Pig"="#E588A6",
                                  "Ruminant"="#0AC5AD","Human"="#3FAAE2"))+
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_jitter(aes(color=Sources), size=2, alpha=0.7, width = 0.2)+
   geom_signif(
    comparisons = comparisons, test = "wilcox.test",
    map_signif_level = FALSE, textsize = 4,
    step_increase = 0.1,color="black"
  ) +
   ylab("crAssphage GC (%)")+
   xlab("")+
   theme(legend.title=element_blank())+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## GC content of *Prevotella* spp.
```{r,warning=FALSE,message=FALSE}
#bacteria_gc
theme_set(theme_llm(base_size = 16,legend = "none"))
GC_bac <- read.table("../Pre-processed_Files/prevotella_gc.txt",header = T)
GC_bac$Sources <- factor(GC_bac$Sources,levels = c("Chicken", "Pig","Ruminant", 
                                                   "Human"))
library(ggsignif)
comparisons <- list(c("Pig", "Ruminant"),c("Chicken", "Pig"), 
                    c("Chicken", "Ruminant"),c("Chicken", "Human"))
ggplot(GC_bac, aes(x=Sources, y=GC, color=Sources)) +
  geom_boxplot(size=0.5) +
   scale_color_manual(values  = c("Chicken"="#FFBC00","Pig"="#E588A6",
                                  "Ruminant"="#0AC5AD","Human"="#3FAAE2"))+
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_jitter(aes(color=Sources), size=2, alpha=0.7, width = 0.2)+
   geom_signif(
    comparisons = comparisons, test = "wilcox.test",
    map_signif_level = FALSE, textsize = 4,
    step_increase = 0.1,color="black"
  ) +
   ylab(expression(paste(italic("Prevotella "), "GC (%)",sep = "")))+
   xlab("")+
   theme(legend.title=element_blank())+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
```

## Heatmap of crAss-like phages relative abundance 
```{r,warning=FALSE,message=FALSE}
chick_phage_all_tpm_df_crass_feces <-read.csv("../Pre-processed_Files/chick_phage_crass.csv",
                                              row.names = 1)

chick_geo_group <- read.table("../Pre-processed_Files/chick_geo_group.txt",
                              header = T,sep = "\t")

chick_geo_group <- column_to_rownames(chick_geo_group,"id")

chick_geo_group <- filter(chick_geo_group,ource=="feces"|ource=="cecum")

annotation_col = chick_geo_group["continent"]
chick_phage_all_tpm_df_crass_feces <- chick_phage_all_tpm_df_crass_feces[,rownames(chick_geo_group)]
rownames(chick_phage_all_tpm_df_crass_feces) <- gsub(pattern = "chick",
                                                     replacement = "chicken",
                                                     x = rownames(chick_phage_all_tpm_df_crass_feces))
library(RColorBrewer)
library(pheatmap,lib.loc = "/home/lsx/R/x86_64-pc-linux-gnu-library/4.4/")
pheatmap::pheatmap(log(chick_phage_all_tpm_df_crass_feces+1),
                   show_rownames = T,
                   show_colnames = F,
                   annotation_col = annotation_col,
                   treeheight_row = 10,
                   treeheight_col = 10,
                   legend = F,
                   breaks = seq(0,12,by=0.12),
                   annotation_legend = T,
                   annotation_colors = list(continent = c("Asia" = "#FB8072",
                                                          "Europe"= "#BEBADA",
                                                          "North America"="#FDB462",
                                                          "Oceania"="#80B1D3")),
                   color = colorRampPalette(colors = c("grey",
                                                       "#fffbdf",
                                                       "orange",
                                                       "red"))(100))


chicken_crass_prevotella <- filter(chick_phage_all_tpm_df_crass_feces,rownames(chick_phage_all_tpm_df_crass_feces) %in% c("chicken_phage_4603","chicken_phage_5252","chicken_phage_3894","chicken_phage_4328","chicken_phage_4126","chicken_phage_4414","chicken_phage_4605","chicken_phage_4020","chicken_phage_5812","chicken_phage_3336","chicken_phage_3600","chicken_phage_5952"))

chicken_crass_prevotella_abundance <- colSums(chicken_crass_prevotella)

chicken_crass_bacteroides <- filter(chick_phage_all_tpm_df_crass_feces,rownames(chick_phage_all_tpm_df_crass_feces) %in% c("chicken_phage_3394","chicken_phage_5709","chicken_phage_3688","chicken_phage_3418","chicken_phage_3104","chicken_phage_5699"))

chicken_crass_bacteroides_abundance <- colSums(chicken_crass_bacteroides)                            
```

```{r,warning=FALSE,message=FALSE}
pig_phage_all_tpm_df_crass <- read.csv("../Pre-processed_Files/pig_phage_crass.csv",row.names = 1)

pig_geo_group <- read.table("../Pre-processed_Files/pig_geo_group.txt",
                            header = T,sep = "\t")
pig_geo_group <- column_to_rownames(pig_geo_group,"Run")
annotation_col = pig_geo_group["continent"]
library(RColorBrewer)
pheatmap::pheatmap(log(pig_phage_all_tpm_df_crass+1),
                   show_rownames = T,
                   show_colnames = F,
                   annotation_col = annotation_col,
                   cluster_rows = TRUE,
                   cluster_cols = TRUE, 
                   treeheight_row = 10,
                   treeheight_col = 10,
                   legend = F,
                   annotation_legend = TRUE,
                   annotation_colors = list(continent = c("Asia" = "#FB8072",
                                                          "Europe"= "#BEBADA",
                                                          "North America"="#FDB462",
                                                          "Oceania"="#80B1D3")),
                   color = colorRampPalette(colors = c("grey",
                                                       "#fffbdf",
                                                       "orange",
                                                       "red"))(100))

```

```{r,warning=FALSE,message=FALSE}
cow_phage_all_tpm_df_crass <- read.csv("../Pre-processed_Files/p_cow_crass.heatmap.csv",row.names = 1)
cow_geo_group <- read.table("../Pre-processed_Files/cow_geo_group.txt",
                            header = T,sep = "\t")
cow_geo_group <- column_to_rownames(cow_geo_group,"id")
rownames(cow_phage_all_tpm_df_crass) <- rownames(cow_phage_all_tpm_df_crass) %>% gsub("cow","ruminant",.)

annotation_col = cow_geo_group["continent"]
library(RColorBrewer)

pheatmap::pheatmap(log(cow_phage_all_tpm_df_crass+1),
                   show_rownames = T,
                   show_colnames = F,
                   annotation_col = annotation_col,
                   treeheight_row = 10,
                   treeheight_col = 10,
                   legend = F,
                   annotation_legend =T,
                   breaks = seq(0,12,by=0.12),
                   annotation_colors = list(continent = c("Asia" = "#FB8072",
                                                          "Europe"= "#BEBADA",
                                                          "North America"="#FDB462",
                                                          "Oceania"="#80B1D3")),
                   color = colorRampPalette(colors = c("grey","#fffbdf","orange","red"))(100))

cow_crass_prevotella <- filter(cow_phage_all_tpm_df_crass,rownames(cow_phage_all_tpm_df_crass) %in% c("ruminant_phage_05510","ruminant_phage_02792","ruminant_phage_26550","ruminant_phage_10648","ruminant_phage_26670","ruminant_phage_27233","ruminant_phage_15105","ruminant_phage_05441","ruminant_phage_28764","ruminant_phage_08269","ruminant_phage_31701","ruminant_phage_32379","ruminant_phage_15569","ruminant_phage_23344","ruminant_phage_03400","ruminant_phage_00888","ruminant_phage_32242","ruminant_phage_11818","ruminant_phage_07141","ruminant_phage_28412","ruminant_phage_12024","ruminant_phage_32506","ruminant_phage_17518","ruminant_phage_06156","ruminant_phage_18784","ruminant_phage_16979","ruminant_phage_27337","ruminant_phage_08542","ruminant_phage_28423","ruminant_phage_29855","ruminant_phage_29620","ruminant_phage_29612","ruminant_phage_29922","ruminant_phage_29859","ruminant_phage_06886","ruminant_phage_06665","ruminant_phage_29722","ruminant_phage_30575","ruminant_phage_12511","ruminant_phage_29998","ruminant_phage_30543","ruminant_phage_29841","ruminant_phage_30134","ruminant_phage_30194","ruminant_phage_29741"))

cow_crass_prevotella_abundance <- colSums(cow_crass_prevotella)

cow_crass_bacteroides <- filter(cow_phage_all_tpm_df_crass,rownames(cow_phage_all_tpm_df_crass) %in% c("ruminant_phage_32242","ruminant_phage_08542"))

cow_crass_bacteroides_abundance <- colSums(cow_crass_bacteroides)
```

## Box plot showing the relative abundance of crAss-like phages within different animals and bacterial hosts

```{r,warning=FALSE,message=FALSE}
chicken_crass_bacteroides_abundance_df <- as.data.frame(chicken_crass_bacteroides_abundance)
chicken_crass_bacteroides_abundance_df$source="Chicken"

cow_crass_bacteroides_abundance_df <- as.data.frame(cow_crass_bacteroides_abundance)
cow_crass_bacteroides_abundance_df$source="Ruminant"

colnames(cow_crass_bacteroides_abundance_df) <- c("Abundance","source")
colnames(chicken_crass_bacteroides_abundance_df) <- c("Abundance","source")
chick_ruminant_crass_bacteroides <- rbind(chicken_crass_bacteroides_abundance_df,
                                          cow_crass_bacteroides_abundance_df)

library(ggsignif)
theme_set(theme_llm(base_size = 16,legend = "none"))
#comparisons01=list(c("Chicken","Ruminant"))
ggplot(chick_ruminant_crass_bacteroides,aes(x=source,y=Abundance,fill=source))+ 
  geom_boxplot()+
  #scale_y_log10()+
  ylab("Relative Abundance of Bacteroides crAss-like Phages")+
  stat_boxplot(geom ='errorbar', width = 0.3)+
  scale_fill_manual(values  = c("Chicken"="#FFBC00","Ruminant"="#0AC5AD"))+
  geom_signif(comparisons = list(c("Chicken","Ruminant")), 
            test = "wilcox.test", 
            textsize = 6, 
            tip_length = 0.01)+
  ylim(10,300)

chicken_crass_prevotella_abundance_df <- as.data.frame(chicken_crass_prevotella_abundance)
chicken_crass_prevotella_abundance_df$source="Chicken"

cow_crass_prevotella_abundance_df <- as.data.frame(cow_crass_prevotella_abundance)
cow_crass_prevotella_abundance_df$source="Ruminant"

colnames(cow_crass_prevotella_abundance_df) <- c("Abundance","source")
colnames(chicken_crass_prevotella_abundance_df) <- c("Abundance","source")
chick_ruminant_crass_prevotella <- rbind(chicken_crass_prevotella_abundance_df,
                                         cow_crass_prevotella_abundance_df)

ggplot(chick_ruminant_crass_prevotella,aes(x=source,y=Abundance,fill=source))+ 
  geom_boxplot()+
  #scale_y_log10()+
  ylab("Relative Abundance of Prevotela crAss-like Phages")+
  stat_boxplot(geom ='errorbar', width = 0.3)+
  ylim(10,10000)+
  scale_fill_manual(values  = c("Chicken"="#FFBC00","Ruminant"="#0AC5AD"))+
  geom_signif(comparisons = list(c("Chicken","Ruminant")), 
            test = "wilcox.test", 
            textsize = 6, 
            tip_length = 0.01)
```

## Prevalence of specific crAss-like phages in chickens and ruminants

```{r,warning=FALSE,message=FALSE}
prevalence_data <- data.frame(
source = rep(c("Ruminant","Chicken"), each = 2),
host = rep(c("Prevotella", "Bacteroides"), times = 2),
value = c(0.9017758,0.3701443,0.8820116,0.6073501)
)
theme_set(theme_llm(base_size = 16,legend = "right"))
ggplot(prevalence_data, aes(x = host, y = value*100, fill = source))+
  geom_bar(stat = "identity", position = "dodge")+
  scale_fill_manual(values  = c("Chicken"="#FFBC00","Ruminant"="#0AC5AD"))+
  ylab("Prevalence (%)")+xlab("")+theme(aspect.ratio=1)+ylim(0,100)

```
## Box plot showing the relative abundance of crAss-like phages in goat, sheep, and dairy cattle
```{r,warning=FALSE,message=FALSE}
goat_sheep_cattle <- read.csv("../Pre-processed_Files/goat_sheep_cattle_crass.csv",header = T)

goat_sheep_cattle$source <- factor(goat_sheep_cattle$source,levels = c("Goat",
                                                                       "Sheep",
                                                                       "Dairy_cattle"))
library(ggsignif)
theme_set(theme_llm(base_size = 16,legend = "none"))
comparisons3=list(c("Goat","Sheep"),c("Sheep","Dairy_cattle"),c("Goat","Dairy_cattle"))
ggplot(goat_sheep_cattle,aes(x=source,y=abundance,fill=source))+ 
  geom_boxplot()+
 ylim(0,350)+
  ylab("TPM")+ggsci::scale_fill_npg()+
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_signif(comparisons = list(c("Goat","Sheep")), 
            test = "wilcox.test", 
            textsize = 6, 
            tip_length = 0.01,
            y_position = c(150,200))+
  geom_signif(comparisons = list(c("Sheep","Dairy_cattle")), 
              test = "wilcox.test", 
              textsize = 6, 
              tip_length = 0.01,
              y_position = c(200,250))+
  geom_signif(comparisons = list(c("Goat","Dairy_cattle")), 
              test = "wilcox.test", 
              textsize = 6, 
              tip_length = 0.01,
              y_position = c(250,300))
```


## Box plot showing the relative abundance of crAss-like phages between wild boars, Duroc pigs and Licha pigs
```{r,warning=FALSE,message=FALSE}
## different breed
pig_breed_group <- read.table("../Pre-processed_Files/pig_breed_group.txt",header = T)
pig_PA <- read.csv("../Pre-processed_Files/pig_crass_in_every_country.csv")
pig_breed_crass_rate <- left_join(pig_breed_group,pig_PA,by="id")
library(ggsignif)
theme_set(theme_llm(base_size = 16,legend = "none"))
pig_breed_crass_rate$Breed <- factor(pig_breed_crass_rate$Breed,
                                     levels = c("Wild-Boar","Duroc","Licha"))
ggplot(pig_breed_crass_rate,aes(x=Breed,y=sum,fill=Breed))+ 
  geom_boxplot()+
 ylim(0,1800)+
  ylab("TPM")+
  stat_boxplot(geom ='errorbar', width = 0.3)+
geom_signif(comparisons = list(c("Duroc", "Licha")), 
            test = "wilcox.test", 
            textsize = 6, 
            tip_length = 0.01,
            y_position = c(800,900))+
  geom_signif(comparisons = list(c("Duroc", "Wild-Boar")), 
              test = "wilcox.test", 
              textsize = 6, 
              tip_length = 0.01,
              y_position = c(500,600))+
  geom_signif(comparisons = list(c("Licha", "Wild-Boar")), 
              test = "wilcox.test", 
              textsize = 6, 
              tip_length = 0.01,
              y_position = c(300,400))+
  ggsci::scale_fill_npg()

```

### sessionInfo
```{r}
sessionInfo()
```