---
title: "DAGPD: Coevolution and coadaptation with Prevotella and animals"
author: "Yuxuan Huang"
format: 
  html:
    embed-resources: true
---

# Coevolution and coadaptation with Prevotella and animals

## GC content of crAss-like phages alpha group

```{r,warning=FALSE,message=FALSE}
# setwd("~/Downloads/DAGPD/cell/Scripts")
library(tidyverse)
source("./theme_llm.R")
theme_set(theme_llm(base_size = 16,legend = "none"))
# alpha group

GC <- read.table("../Pre-processed_Files/Alpha_GC.txt",header = T)
GC$Sources <- factor(GC$Sources,levels = c("Chicken", "Pig","Ruminant", "Human"))
library(ggsignif)
comparisons <- list(c("Pig", "Ruminant"),c("Ruminant", "Human"),c("Pig","Human"),c("Chicken", "Pig"), c("Chicken", "Ruminant"),c("Chicken", "Human"))
ggplot(GC, aes(x=Sources, y=GC, color=Sources)) +
  geom_boxplot(size=0.5) +
   scale_color_manual(values  = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD","Human"="#3FAAE2"))+
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_jitter(aes(color=Sources), size=2, alpha=0.7, width = 0.2)+
   geom_signif(
    comparisons = comparisons, test = "t.test",
    map_signif_level = F, textsize = 4,
    step_increase = 0.1,color="black"
  ) +
   ylab("crAssphage GC (%)")+
   xlab("")+
   theme(legend.title=element_blank())+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## GC content of crAss-like phages beta group
```{r,warning=FALSE,message=FALSE}
# beta group
theme_set(theme_llm(base_size = 16,legend = "none"))
GC <- read.csv("../Pre-processed_Files/crass_beta_gc.csv",header = T)
GC$Sources <- factor(GC$Sources,levels = c("Chicken", "Pig","Ruminant", "Human"))
library(ggsignif)
comparisons <- list(c("Pig", "Ruminant"),c("Ruminant", "Human"),c("Pig","Human"),c("Chicken", "Pig"), c("Chicken", "Ruminant"),c("Chicken", "Human"))
ggplot(GC, aes(x=Sources, y=GC, color=Sources)) +
  geom_boxplot(size=0.5) +
   scale_color_manual(values  = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD","Human"="#3FAAE2"))+
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_jitter(aes(color=Sources), size=2, alpha=0.7, width = 0.2)+
   geom_signif(
    comparisons = comparisons, test = "t.test",
    map_signif_level = FALSE, textsize = 4,
    step_increase = 0.1,color="black"
  ) +
   ylab("crAssphage GC (%)")+
   xlab("")+
   theme(legend.title=element_blank())+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## GC content of *Prevotella* spp.
```{r,warning=FALSE,message=FALSE}
#bacteria_gc
theme_set(theme_llm(base_size = 16,legend = "none"))
GC_bac <- read.table("../Pre-processed_Files/prevotella_gc.txt",header = T)
GC_bac$Sources <- factor(GC_bac$Sources,levels = c("Chicken", "Pig","Ruminant", "Human"))
library(ggsignif)
comparisons <- list(c("Pig", "Ruminant"),c("Chicken", "Pig"), c("Chicken", "Ruminant"),c("Chicken", "Human"))
ggplot(GC_bac, aes(x=Sources, y=GC, color=Sources)) +
  geom_boxplot(size=0.5) +
   scale_color_manual(values  = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD","Human"="#3FAAE2"))+
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_jitter(aes(color=Sources), size=2, alpha=0.7, width = 0.2)+
   geom_signif(
    comparisons = comparisons, test = "t.test",
    map_signif_level = FALSE, textsize = 4,
    step_increase = 0.1,color="black"
  ) +
   ylab(expression(paste(italic("Prevotella "), "GC (%)",sep = "")))+
   xlab("")+
   theme(legend.title=element_blank())+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
```

## Heatmap of crAss-like phages relative abundance 
```{r,warning=FALSE,message=FALSE}
chick_phage_all_tpm_df_crass_feces <-read.csv("../Pre-processed_Files/chick_phage_crass.csv",row.names = 1)

chick_geo_group <- read.table("../Pre-processed_Files/chick_geo_group.txt",header = T,sep = "\t")

chick_geo_group <- column_to_rownames(chick_geo_group,"id")

chick_geo_group <- filter(chick_geo_group,ource=="feces"|ource=="cecum")

annotation_col = chick_geo_group["continent"]
chick_phage_all_tpm_df_crass_feces <- chick_phage_all_tpm_df_crass_feces[,rownames(chick_geo_group)]
rownames(chick_phage_all_tpm_df_crass_feces) <- gsub(pattern = "chick",replacement = "chicken",x = rownames(chick_phage_all_tpm_df_crass_feces))
library(RColorBrewer)
pheatmap::pheatmap(log(chick_phage_all_tpm_df_crass_feces+1),
                   show_rownames = T,
                   show_colnames = F,
                   annotation_col = annotation_col,
                   treeheight_row = 10,
                   treeheight_col = 10,
                   legend = F,
                   breaks = seq(0,12,by=0.12),
                   annotation_legend = T,
                   annotation_colors = list(continent = c("Asia" = "#FB8072",
                                                          "Europe"= "#BEBADA",
                                                          "North America"="#FDB462",
                                                          "Oceania"="#80B1D3")),
                   color = colorRampPalette(colors = c("grey","#fffbdf","orange","red"))(100))

```

```{r,warning=FALSE,message=FALSE}
pig_phage_all_tpm_df_crass <- read.csv("../Pre-processed_Files/pig_phage_crass.csv",row.names = 1)

pig_geo_group <- read.table("../Pre-processed_Files/pig_geo_group.txt",header = T,sep = "\t")
pig_geo_group <- column_to_rownames(pig_geo_group,"Run")
annotation_col = pig_geo_group["continent"]
library(RColorBrewer)
pheatmap::pheatmap(log(pig_phage_all_tpm_df_crass+1),
                   show_rownames = T,
                   show_colnames = F,
                   annotation_col = annotation_col,
                   cluster_rows = TRUE,
                   cluster_cols = TRUE, 
                   treeheight_row = 10,
                   treeheight_col = 10,
                   legend = F,
                   annotation_legend = TRUE,
                   annotation_colors = list(continent = c("Asia" = "#FB8072",
                                                          "Europe"= "#BEBADA",
                                                          "North America"="#FDB462",
                                                          "Oceania"="#80B1D3")),
                   color = colorRampPalette(colors = c("grey","#fffbdf","orange","red"))(100))

```

```{r,warning=FALSE,message=FALSE}
cow_phage_all_tpm_df_crass <- read.csv("../Pre-processed_Files/p_cow_crass.heatmap.csv",row.names = 1)
cow_geo_group <- read.table("../Pre-processed_Files/cow_geo_group.txt",header = T,sep = "\t")
cow_geo_group <- column_to_rownames(cow_geo_group,"id")
rownames(cow_phage_all_tpm_df_crass) <- rownames(cow_phage_all_tpm_df_crass) %>% gsub("cow","ruminant",.)

annotation_col = cow_geo_group["continent"]
library(RColorBrewer)

pheatmap::pheatmap(log(cow_phage_all_tpm_df_crass+1),
                   show_rownames = T,
                   show_colnames = F,
                   annotation_col = annotation_col,
                   treeheight_row = 10,
                   treeheight_col = 10,
                   legend = F,
                   annotation_legend =T,
                   breaks = seq(0,12,by=0.12),
                   annotation_colors = list(continent = c("Asia" = "#FB8072",
                                                          "Europe"= "#BEBADA",
                                                          "North America"="#FDB462",
                                                          "Oceania"="#80B1D3")),
                   color = colorRampPalette(colors = c("grey","#fffbdf","orange","red"))(100))


```


## Box plot showing the relative abundance of crAss-like phages in goat, sheep, and dairy cattle
```{r,warning=FALSE,message=FALSE}
goat_sheep_cattle <- read.csv("../Pre-processed_Files/goat_sheep_cattle_crass.csv",header = T)

goat_sheep_cattle$source <- factor(goat_sheep_cattle$source,levels = c("Goat","Sheep","Dairy_cattle"))
library(ggsignif)
theme_set(theme_llm(base_size = 16,legend = "none"))
comparisons3=list(c("Goat","Sheep"),c("Sheep","Dairy_cattle"),c("Goat","Dairy_cattle"))
ggplot(goat_sheep_cattle,aes(x=source,y=abundance,fill=source))+ 
  geom_boxplot()+
 ylim(0,350)+
  ylab("TPM")+ggsci::scale_fill_npg()+
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_signif(comparisons = list(c("Goat","Sheep")), 
            test = "t.test", 
            textsize = 6, 
            tip_length = 0.01,
            y_position = c(150,200))+
  geom_signif(comparisons = list(c("Sheep","Dairy_cattle")), 
              test = "t.test", 
              textsize = 6, 
              tip_length = 0.01,
              y_position = c(200,250))+
  geom_signif(comparisons = list(c("Goat","Dairy_cattle")), 
              test = "t.test", 
              textsize = 6, 
              tip_length = 0.01,
              y_position = c(250,300))
```


## Box plot showing the relative abundance of crAss-like phages between wild boars, Duroc pigs and Licha pigs
```{r,warning=FALSE,message=FALSE}
## different breed
pig_breed_group <- read.table("../Pre-processed_Files/pig_breed_group.txt",header = T)
pig_PA <- read.csv("../Pre-processed_Files/pig_crass_in_every_country.csv")
pig_breed_crass_rate <- left_join(pig_breed_group,pig_PA,by="id")
library(ggsignif)
theme_set(theme_llm(base_size = 16,legend = "none"))
pig_breed_crass_rate$Breed <- factor(pig_breed_crass_rate$Breed,levels = c("Wild-Boar","Duroc","Licha"))
ggplot(pig_breed_crass_rate,aes(x=Breed,y=sum,fill=Breed))+ 
  geom_boxplot()+
 ylim(0,1800)+
  ylab("TPM")+
geom_signif(comparisons = list(c("Duroc", "Licha")), 
            test = "t.test", 
            textsize = 6, 
            tip_length = 0.01,
            y_position = c(800,900))+
  geom_signif(comparisons = list(c("Duroc", "Wild-Boar")), 
              test = "t.test", 
              textsize = 6, 
              tip_length = 0.01,
              y_position = c(500,600))+
  geom_signif(comparisons = list(c("Licha", "Wild-Boar")), 
              test = "t.test", 
              textsize = 6, 
              tip_length = 0.01,
              y_position = c(300,400))+
  ggsci::scale_fill_npg()

```

### sessionInfo
```{r}
sessionInfo()
```
