---
title: "DAGPD: Host assignment and host-linked phage abundance for domestic animal gut phage"
author: "Yuxuan Huang"
format: 
  html:
    embed-resources: true
---

# Host assignment and host-linked phage abundance for domestic animal gut phage

```{r,message=FALSE,warning=FALSE}
# setwd("~/Downloads/DAGPD/cell/Scripts")
library(tidyverse)
source("./theme_llm.R")

host <- read.table("../Pre-processed_Files/hosts1003.txt",header = T,sep="\t")
host$Hosts <- factor(host$Hosts,levels = unique(host$Hosts))
host$Species <- factor(host$Species,levels = c("sum","Chicken", "Pig", "Ruminant","Human"))
host$Phylum <- factor(host$Phylum,levels = unique(host$Phylum))
#p2 <- ggplot()  + 
# geom_line(mapping = aes(x = data2$VC, y = data2$Genomes,group=1),color="orange")

theme_set(theme_llm(base_size = 12,margin=FALSE))
color_mapping <- c("Firmicutes"="#AFF05B","Bacteroidota"="#6E40AA","Proteobacteria"="#00BFFF","Spirochaetota"="#FFD92F","Pseudomonadota"="#E5C494","Actinobacteriota"="#FC8D62","Campylobacterota"="#8DA0CB","Fibrobacterota"="#E78AC3")
ggplot()  + 
  geom_col(data= filter(host,Species!="sum"),
           mapping = aes(x=Hosts,y=number ,fill=Species),
           alpha=0.7) +
  labs(fill="Sources")+
  scale_color_manual(
    values = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD","Human"="#3FAAE2"),
    breaks = c("Chicken", "Pig", "Ruminant","Human"),
  aesthetics = c("fill"))+
  geom_point(data= filter(host,Species=="sum"),
             group=1,mapping=aes(x=Hosts,y=number+0.25,colour=Phylum),size = 2) +
  scale_color_manual(values = color_mapping)+
  labs(y = "Genomes(%)",x="Genus of Hosts")+
  theme(aspect.ratio=0.32)+
  #ggsci::scale_fill_npg()+
  #scale_color_manual(values  = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD"),aesthetics = c("fill"))+
  theme(legend.position = c(0.85,0.53))+
  theme(legend.text = element_text(size = 8))+
  theme(legend.key.size = unit(1,"mm"))+
  theme(legend.margin = margin(0.1,0.1,0.1,0.1,'mm'))+
  theme(legend.title = element_text(size = 8))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Principal Co-ordinates Analysis (PCoA) plot of Bray-Curtis dissimilarity matrix calculated from TPM value of common VCs
```{r,warning=FALSE,message=FALSE}
chick_pig_cow_pcoa <- read.csv("../Pre-processed_Files/chick_pig_cow_poca.point.csv",header = T)
chick_pig_cow_pcoa <- column_to_rownames(chick_pig_cow_pcoa,"X")

theme_set(theme_llm(base_size = 16,legend = "right"))
ggplot(chick_pig_cow_pcoa, aes(x=V1, y=V2, colour=Source))+geom_point(alpha=0.7, size=1)+
  scale_color_manual(
    values = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD"))+
  stat_ellipse()+
  xlab("PC1 (14.8%)")+
  ylab("PC2 (10.6%)")

```

## Box plot showing the GC content among phage genomes at different host phylum (left) and different phage family (right)

```{r,warning=FALSE,message=FALSE}
gc_content <- read.csv("../Pre-processed_Files/gc_content.csv",header = T)
library(ggsignif)

comparisons <- list(c("Bacteroidota", "Firmicutes"),c("Firmicutes", "Proteobacteria"),c("Bacteroidota","Proteobacteria"))
theme_set(theme_llm(base_size = 16,legend = "none"))
p1 <- ggplot(filter(gc_content,Category=="Phylum"), aes(x = Group, y = GC,color=Group)) +
    geom_boxplot(size=0.5)+ 
  scale_color_manual(values  = c("Bacteroidota"="#6E40AA","Firmicutes"="#AFF05B","Proteobacteria"="#00BFFF"))+
   
      stat_boxplot(geom ='errorbar', width = 0.3) +

 geom_jitter(aes(color=Group), size=2, alpha=0.7, width = 0.2)+
  geom_signif(
    comparisons = comparisons, test = "t.test",
    map_signif_level = TRUE, textsize = 4,
    step_increase = 0.1,color="black"
  )+
  theme(aspect.ratio=2)+
  ylab("Phage GC content (%)")+
   xlab("")+
   theme(legend.title=element_blank())+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
    
## family group
comparisons2 <- list(c("Herelleviridae", "Salasmaviridae"),c("Salasmaviridae","Schitoviridae"),c("Schitoviridae","Drexlerviridae"),c("Herelleviridae","Schitoviridae"),c("Salasmaviridae","Drexlerviridae"),c("Herelleviridae","Drexlerviridae"))
theme_set(theme_llm(base_size = 16,legend = "none"))
gc_content$Group <- factor(gc_content$Group,levels = c("Herelleviridae", "Salasmaviridae","Schitoviridae", "Drexlerviridae"))
p2 <- ggplot(filter(gc_content,Category=="Taxonomy"), aes(x = Group, y = GC,color=Group)) +
    geom_boxplot(size=0.5)+ 
  scale_color_manual(values  = c("Drexlerviridae"="#FCCDE5","Herelleviridae"="#E392FE","Salasmaviridae"="#F5EC00","Schitoviridae"="#8DD3C7"))+
   
      stat_boxplot(geom ='errorbar', width = 0.3) +

  geom_jitter(aes(color=Group), size=2, alpha=0.7, width = 0.2)+
  geom_signif(
    comparisons = comparisons2, test = "t.test",
    map_signif_level = TRUE, textsize = 4,
    step_increase = 0.03,color="black"
  )+
  theme(aspect.ratio=2)+
  ylim(c(25,70))+
  ylab("Phage GC content (%)")+
   xlab("")+
   theme(legend.title=element_blank())+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
library(patchwork)
p1|p2
```


## Box plot showing the mapping rate and ratio of phages and bacteria in different domestic animals.
```{r,warning=FALSE,message=FALSE}
mappingrate <- read.table("../Pre-processed_Files/mappingrate.txt",header = T)
mappingrate$Category <- factor(mappingrate$Category, levels = c("Phage", "Bacteria", "Ratio"))

theme_set(theme_llm(base_size = 16,legend = "top",margin=FALSE))

# grouped boxplot
ggplot(mappingrate, aes(x=Species, y=Rate, fill=Category)) + 
  geom_boxplot(lwd=0.2,outlier.size = 0.2,outlier.color = "grey")+
  scale_fill_discrete(limits = c("phage", "Bacteria", "Ratio"))+
  labs(x = "Sources",y="Mapping Rate")+
  theme(aspect.ratio=1)+
  ggsci::scale_fill_npg()+
  theme(legend.title=element_blank())

```

## Significant Pearson correlation between relative abundances of representative phages and their hosts in different domestic animals


```{r,warning=FALSE,message=FALSE}
chick_pig_cow_cor <- read.csv("../Pre-processed_Files/chick_pig_cow_cor.csv",header = T)
ggplot(chick_pig_cow_cor, aes(x = log(Host,2), y = log(Phage,2), color = Source)) +
  geom_point(size=0.5,alpha = 0.5) +  
  ylim(0,20)+
  geom_smooth(method = "lm", se = T,size=0.5)+theme(aspect.ratio=1)+
  scale_color_manual(values  = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD"))+
  theme(aspect.ratio=1)

```


## NMDS analysis of the core phage communities
```{r,message=FALSE,warning=FALSE}
chick_pig_cow_nmds <- read.csv("../Pre-processed_Files//NMDSpoints1_chick_pig_cow.csv",header = T)
chick_pig_cow_nmds <- column_to_rownames(chick_pig_cow_nmds,"X")
ggplot(chick_pig_cow_nmds, aes(x=MDS1, y=MDS2, colour=Source))+geom_point(alpha=0.7, size=1)+
  scale_color_manual(
    values = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD"))+
  stat_ellipse()

```

## PCoA analysis of the core bacterial communities 
```{r,message=FALSE,warning=FALSE}
chick_pig_cow_bac_pcoa <- read.csv("../Pre-processed_Files/pcoa_bac_points.csv",header = T)

chick_pig_cow_bac_pcoa <- column_to_rownames(chick_pig_cow_bac_pcoa,"X")
ggplot(chick_pig_cow_bac_pcoa, aes(x=V1, y=V2, colour=Source))+geom_point(alpha=0.7, size=1)+
  scale_color_manual(
    values = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD"))+
  stat_ellipse()+
  xlab("PCoA1 (20.66%)")+
  ylab("PCoA2 (16.57%)")
```

## NMDS analysis of the core bacterial communities 
```{r,message=FALSE,warning=FALSE}
chick_pig_cow_bac_nmds <- read.csv("../Pre-processed_Files/bacNMDSpoints1.csv",header = T)

#chick_pig_cow_bac_nmds <- column_to_rownames(chick_pig_cow_bac_nmds,"X")
ggplot(chick_pig_cow_bac_nmds, aes(x=MDS1, y=MDS2, colour=Source))+geom_point(alpha=0.7, size=1)+
  scale_color_manual(
    values = c("Chicken"="#FFBC00","Pig"="#E588A6","Ruminant"="#0AC5AD"))+
  stat_ellipse()

```
